An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-17 EconGrowth-R-Functions.Rnw
Now including ZA flag. Now including data from US Single Factor fits. ZA single factor fits not yet working.

#############################
# Creates a graph with confidene intervals for the single-factor model for the given factor
##
createSFParamsGraph <- function(factor){
  # Set the name of the 2nd graph (m)
  factorLabel <- paste("$", tolower(factor), "$", sep="")
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda or m), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, singleFactorCountryRowsForParamsGraph, factor=factor))
  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable,
                   centers = value, #identifies where the dots should be placed
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(2,1), #2 column, 1 row
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   strip = strip.custom(bg="white", factor.levels=c("$\\lambda$", factorLabel)),
                   ylim = list(c(-0.05, 0.1), c(-0.5, 1.5)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.0, 0.05, 0.1),   #y tick mark for lambda
                                                c(0.0, 0.5, 1.0)
                                                )
                                        )
                                 )
                   )

>>>>

#############################
# Creates a graph with confidene intervals for the single-factor model for the given factor
##
createSFParamsGraph <- function(factor){
  # Figure out the exponent for the desired factor
  if (factor == "K"){
    exponentParameter <- "$\\alpha$"
  } else if (factor == "L"){
    exponentParameter <- "$\\beta$"
  } else if ((factor == "Q") || (factor == "X") || (factor == "U")){
    exponentParameter <- "$\\gamma$"
  }
  # Create a data table with the following columns:
  # country abbrev, parameter (lambda or m), -95% CI, value, +95% CI
  dataTable <- do.call("rbind", lapply(countryAbbrevs, singleFactorCountryRowsForParamsGraph, factor=factor))
  graph <- segplot(country ~ upperCI + lowerCI | parameter, 
                   data = dataTable,
                   centers = value, #identifies where the dots should be placed
                   draw.bands = FALSE, #provides nicer error bars
                   horizontal = FALSE, #makes error bars vertical and puts the countries in the x axis
                   layout = c(2,1), #2 column, 1 row
                   col = "black", #Sets line color to black
                   lwd = 1, #Sets line width to 1.0
                   strip = strip.custom(bg="white", factor.levels=c("$\\lambda$", exponentParameter)),
                   ylim = list(c(-0.05, 0.1), c(-0.5, 1.5)), #y axis limits
                   scales = list(cex=scaleTextSize, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 x=list(cex=0.75), #reduces text size so that country abbrevs are legible
                                 y=list(relation="free", #allow each axis to be different
                                        at=list(c(0.0, 0.05, 0.1),   #y tick mark for lambda
                                                c(0.0, 0.5, 1.0)
                                                )
                                        )
                                 )
                   )

### 2013-03-17 EconGrowthSeminar.Rnw
Now including ZA flag. Now including data from US Single Factor fits. ZA single factor fits not yet working.

\begin{frame}{Energy Cost Share and Economic Growth}{\citep{Bashmakov:2007ek}} %----------------------------
  \begin{center}
    \includegraphics[height=6.3cm]{figure_other/Figure_Bashmakov_2007.pdf}
  \end{center}
\end{frame}

\begin{frame}{Energy Cost Share and Economic Growth}{\citep{Murphy:2011jh}}  %----------------------------
  \begin{center}
    \includegraphics[height=6.3cm]{figure_other/Figure_Murphy_2011.png}
  \end{center}  
\end{frame}

@@

\begin{frame} %----------------------------
  \begin{center} \huge
    SOURCES OF DATA
  \normalsize \end{center}
\end{frame}

@@

\begin{frame} %----------------------------
  \begin{center} \huge
    ANALYSIS OF GROWTH MODELS
  \normalsize \end{center}
\end{frame}

@@

\begin{frame}{Example Single-Factor Fit} %----------------------------
  Heun complete
\end{frame}

@@

\begin{frame}{US: Single Factor and Cobb-Douglas} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
<<Results_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createSFLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
    \column{2.2in}
<<Results_US_CD, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createCDLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
\end{frame}

@@

>>>>

\begin{frame}{Energy Cost Share}{\citep{Bashmakov:2007ek}} %----------------------------
  \begin{center}
    \includegraphics[height=6.3cm]{figure_other/Figure_Bashmakov_2007.pdf}
  \end{center}
\end{frame}

\begin{frame}{Energy Cost Share and Economic Growth}{\citep{Murphy:2011jh}}  %----------------------------
  \begin{center}
    \includegraphics[height=6.0cm]{figure_other/Figure_Murphy_2011.png}
  \end{center}  
\end{frame}

@@

\begin{frame} %----------------------------
  \begin{center} \huge
    DATA SOURCES
  \normalsize \end{center}
\end{frame}

@@

\begin{frame} %----------------------------
  \begin{center} \huge
    ANALYSIS OF\\
    \vspace{\baselineskip}
    GROWTH MODELS
  \normalsize \end{center}
\end{frame}

@@

\begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_us.pdf} Single-Factor Fit} %----------------------------
<<Results_Table_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
sfUSK <- singleFactorModel("US", "K")
sfUSL <- singleFactorModel("US", "L")
sfUSQ <- singleFactorModel("US", "Q")
sfUSX <- singleFactorModel("US", "X")
sfUSU <- singleFactorModel("US", "U")
@
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
        \vspace{\baselineskip}

        \tiny
          \begin{tabular}{|c|c|c|}\hline
            $f$   & $\lambda$                      & $m$\\
            \hline
            $k$   & \Sexpr{coef(sfUSK)["lambda"]}  & \Sexpr{coef(sfUSK)["m"]}\\
            \hline
            $l$   & \Sexpr{coef(sfUSL)["lambda"]}  & \Sexpr{coef(sfUSL)["m"]}\\
            \hline
            $q$   & \Sexpr{coef(sfUSQ)["lambda"]}  & \Sexpr{coef(sfUSQ)["m"]}\\
            \hline
            $x$   & \Sexpr{coef(sfUSX)["lambda"]}  & \Sexpr{coef(sfUSX)["m"]}\\
            \hline
            $u$   & \Sexpr{coef(sfUSU)["lambda"]}  & \Sexpr{coef(sfUSU)["m"]}\\
            \hline
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createSFLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
\end{frame}

\begin{frame}{\includegraphics[height=0.25in]{figure_other/flag_za.png} Single-Factor Fit} %----------------------------
<<Results_Table_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
sfZAK <- singleFactorModel("ZA", "K")
sfZAL <- singleFactorModel("ZA", "L")
sfZAQ <- singleFactorModel("ZA", "Q")
sfZAX <- singleFactorModel("ZA", "X")
sfZAU <- singleFactorModel("ZA", "U")
@
  \begin{columns}[c]
    \column{2.2in}
%       \begin{center}
%         $y = \mathrm{e}^{\lambda (t-t_0)}f^{m}$
%         \vspace{\baselineskip}
%         
%         \tiny
%           \begin{tabular}{|c|c|c|}\hline
%             $f$   & $\lambda$                      & $m$\\
%             \hline
%             $k$   & \Sexpr{coef(sfZAK)["lambda"]}  & \Sexpr{coef(sfZAK)["m"]}\\
%             \hline
%             $l$   & \Sexpr{coef(sfZAL)["lambda"]}  & \Sexpr{coef(sfZAL)["m"]}\\
%             \hline
%             $q$   & \Sexpr{coef(sfZAQ)["lambda"]}  & \Sexpr{coef(sfZAQ)["m"]}\\
%             \hline
%             $x$   & \Sexpr{coef(sfZAX)["lambda"]}  & \Sexpr{coef(sfZAX)["m"]}\\
%             \hline
%             $u$   & \Sexpr{coef(sfZAU)["lambda"]}  & \Sexpr{coef(sfZAU)["m"]}\\
%             \hline
%           \end{tabular}
%         \normalsize
%       \end{center}
    \column{2.2in}
<<Results_Graph_ZA_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createSFLatticeGraph("ZA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
\end{frame}

@@

\begin{frame}{US: Single Factor and Cobb-Douglas} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
% <<Results_US_SF, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
% createSFLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
% @
    \column{2.2in}
<<Results_US_CD, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight, fig.pos="H">>=
createCDLatticeGraph("US", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
\end{frame}

@@

\begin{frame}{Acknowledgements} %----------------------------
  \begin{itemize}
    \item{Martin de Wit, Stellenbosch University}
    \item{Randall Pruim, Calvin College}
    \item{Loren L. Heun, Western Michigan University}
    \vspace{\baselineskip}
    \item{Jack and Lois Kuipers Applied Mathematics Endowment\\(Reese and Timmer Summer 2012 funding)}
    \item{Calvin College Alumni Association\\(Reese Fall 2012 funding)}
  \end{itemize}  
\end{frame}
